Waving and swaying

### 2013-12-17 Paper/grass.tex
update text file

\section{Introduction}
Rich set of dynamics are exhibited by sea grasses due to its interaction with water flow which can considerably affect the hydrodynamic conditions near the 
sea grasses, resulting change in hydrodynamic conditions can influence number of environmental 
variables and processes relative to unobstructed flow such as transport of heat, water, vapor, sediments, contaminants, dissolved oxygen 
and various other nutrients; plant growth and biomass production (known as thigmomorphogenesis); vegetation and land temperature etc. 
%Driven by such applications number of scientific work regarding flow over canopies has been done over 
%last three decades. One of main development is recognition that major contribution to canopy turbulence arises from coherent eddies
%of canopy scales.\newline
Among many dynamics shown by flexible grasses due to their interaction with water flow, the response to steady current in form of large amplitude coherent oscillations known as
Monami (Ackerman and Okubo 1993) is focus of this paper. Similar phenomenon of large amplitude oscillation of terrestrial canopy in wind have also been observed named as Honami 
(Inoue 1993). Experimental observations (Grizzle 1996) have shown that monami is observed when mean flow speed exceed a certain threshold value. 
\newline
A key feature of flow through submerged vegetation is the existence of strong shear near top of canopy due to
different amount of drag experienced by fluid in and above the canopy, existence of strong shear is manifested by
existence of an inflection point near top of canopy in velocity profile. 
%A crucial feature of a velocity inflection is that it induces a hydrodynamic instability
%process which sets the pattern of coherent eddies propagating over terrestrial and submerged canopies. In response to coherent eddies canopies
%may exhibit synchronous large-amplitude waving (termed as monami in aquatic setting by Ackerman and Okubo[1993], and honami by Inoue (1955) in terrestrial setting). 
%Monami (Honami) can be observed when flow velocity is above certain threshold value, which can depends on flow depth, elasticity of canopy material
%etc.\newline
Inspired by work of Raupach (1996) for the case of terrestrial grass, previous studies (Ghisalberti 2002) of monami 
have been done under assumption that a pattern of coherent eddies propagating downstream is set by a hydrodynamic instability of shear layer 
similar to that of free shear flow, response of sea grass to coherent eddies propagating downstream results in large amplitude synchronous oscillations known as monami. 
Among some of drawback of previous theory includes
(1) While atmospheric flow through terrestrial vegetation seems similar to aquatic flow through marine grass, atmospheric flow is unbounded in vertical direction 
but aquatic flow is not so its not obvious that similar mechanism will hold in aquatic case too (2) The proposed hydrodynamic instability mechanism of shear layer produced 
by presence of grass does not take into account the presence of grass itself (3) It does not predict any threshold condition for waving to happen 
\newline     
In this paper we propose a coupled fluid-canopy model to study dynamics of flexible canopy and flow field. Canopies are represented by a continuous 
medium which is coupled to flow field via a drag term. We did linear stability analysis of flow without ignoring presence of grass via solving 
a generalized orr-sommerfield equation with presence of grass. Our calculation predicts threshold a threshold criteria for waving which is characterized by Reynolds number.
We further show that for flexible marine grass monami is caused by instability of flow above the grass rather than by shear layer. We confirm our analysis by comparing 
prediction from theory with existing data

>>>>

\section{Introduction}
Rich set of dynamics are exhibited by sea grasses due to their interaction with water flow which can considerably affect the hydrodynamic conditions near the 
grass. Resulting change in hydrodynamic conditions relative to unobstructed flow can influence number of environmental 
variables and processes such as transport of sediments, contaminants, dissolved oxygen;
plant growth and biomass production (known as thigmomorphogenesis) etc. 
%Driven by such applications number of scientific work regarding flow over canopies has been done over 
%last three decades. One of main development is recognition that major contribution to canopy turbulence arises from coherent eddies
%of canopy scales.\newline
Among many dynamics shown by flexible grasses due to their interaction with water flow, their response to steady current in form of large amplitude coherent oscillations known as
mo-nami (Ackerman and Okubo 1993) is focus of this paper. Similar phenomenon of large amplitude coherent oscillation of terrestrial canopy in wind have also been observed named as Honami 
(Inoue 1993).
% Experimental observations (Grizzle 1996) have shown that monami is observed when mean flow speed exceed a certain threshold value. 
\newline
A key feature of flow through submerged vegetation is the existence of strong shear near top of canopy due to
different amount of drag experienced by fluid in and above the canopy, existence of strong shear is manifested by
existence of an inflection point in mean velocity near top of canopy. In addition pattern of coherent eddies propagating downstream 
along have also been observed.
%A crucial feature of a velocity inflection is that it induces a hydrodynamic instability
%process which sets the pattern of coherent eddies propagating over terrestrial and submerged canopies. In response to coherent eddies canopies
%may exhibit synchronous large-amplitude waving (termed as monami in aquatic setting by Ackerman and Okubo[1993], and honami by Inoue (1955) in terrestrial setting). 
%Monami (Honami) can be observed when flow velocity is above certain threshold value, which can depends on flow depth, elasticity of canopy material
%etc.
\newline
Inspired by work of Raupach (1996) for the case of terrestrial grass, previous studies (Ghisalberti 2002) of monami 
have been done under assumption that the a pattern of coherent eddies are excited by hydrodynamic instability of shear layer
similar to that of free shear flow. Response of sea grasses to these coherent eddies are exhibited in form of large amplitude oscillations knows as mo-nami.
Several aspect of existing theory remain unclear. The proposed hydrodynamic instability of shear layer layer does not take into account presence of grass whereas shear layer 
was produced due to drag caused by presence vegetation. Since shear layer instability mechanism does not predict any threshold for monami, hence it fails to explain
observations where monami is observed when mean flow speed is above certain value. While atmospheric flow through terrestrial vegetation seems similar to aquatic flow through marine grass, 
atmospheric flow is unbounded in vertical direction.
% but aquatic flow is not so its not obvious that similar mechanism will hold in aquatic case too (2) The proposed hydrodynamic instability mechanism of shear layer produced 
%by presence of grass does not take into account the presence of grass itself (3) It does not predict any threshold condition for waving to happen 
\newline     
In this paper we propose a coupled fluid-canopy model to study dynamics of flexible canopy and flow field. Canopies are represented by a continuous 
medium which is coupled to flow field via a drag term. We did linear stability analysis of flow without ignoring presence of grass via solving 
a generalized orr-sommerfield equation with presence of grass. Our calculation predicts threshold a threshold criteria for waving which is characterized by Reynolds number.
We further show that for flexible marine grass monami is caused by instability of flow above the grass rather than by shear layer. We confirm our analysis by comparing 
prediction from theory with existing data
